---
title: "berryFunctions"
author: "Berry Boessenkool, <berry-b@gmx.de>"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    \\ toc_float: true \\ not possible as of march 2016
vignette: >
  %\VignetteIndexEntry{berryFunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*The R package `berryFunctions`, available at [github.com/brry](https://github.com/brry/berryFunctions) and [CRAN](https://cran.r-project.org/package=berryFunctions), contains my collection of miscellaneous functions. A lot is related to plotting and hydrology (Vignette [Rmd source](https://raw.githubusercontent.com/brry/berryFunctions/master/vignettes/berryFunctions.Rmd)).*


## Package installation
```{r instcran, eval=FALSE}
install.packages("berryFunctions")
library(berryFunctions)
```
To get the most recent development version from github:
```{r instgit, eval=FALSE}
source("https://raw.githubusercontent.com/brry/berryFunctions/master/R/instGit.R")
instGit("brry/berryFunctions")
```

```{r library, echo=FALSE}
library(berryFunctions)
```

[TOC](#header)

## Package highlights

Scatterpoints with third dimension classified into colors: `colPoints`, `colPointsLegend`, `colPointsHist`  
(This uses severall helper functions like `smallPlot`, `classify`, `logSpaced`, `pretty2`, `seqPal`)  
Write text with colored shape underneath: `textField`  
Histogram of data with logarithmic axis: `logHist`, using `logAxis`
```{r colPoints, fig.show='hold', echo=-c(1:2)}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
set.seed(007)
x <- sample(1:87, 150, TRUE);   y <- sample(1:61, 150, TRUE);  z <- diag(volcano[x,y])-95
colPoints(x,y,z,  pch="+", legargs=list(y2=80, title="Elevation"), add=FALSE)
mtext("colPoints, textField", adj=0.05, outer=TRUE, line=-1.2, cex=1.2, font=2)

text(60,30, "unreadable text")
textField(60, 15, "good text", field="round", fill="orange", cex=1.2)

dat <- rbeta(1e4, 2, 80)*100; dat <- dat[dat>0.1]
logHist(dat, col="tan", breaks=50, main="logHist, logAxis")
```

Linear storage cascade (rainfall-runoff modelling): `lsc`, `unitHydrograph`, `superPos`, `nse`, `rmse`  
```{r lsc, fig.show='hold', echo=-1, fig.height=3.5, fig.width=5.5, warning=FALSE}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
# estimate parameters for Unit Hydrograph, plot data and simulation: lsc
QOBS <- dbeta(1:40/40, 3, 10) + rnorm(20,0,0.2) + c(seq(0,1,len=20), rep(1,20))
PREC <- c(1,1,3,4,5,5,4,3,1,1, rep(0,30))
lsc(PREC, QOBS, area=10, main="lsc, unitHydrograph, superPos") # , plotsim=F
```

Quick linear Regression: `linReg`  
Draw circle with given radius: `circle`  
Add transparency to existing colors: `addAlpha`  
Fit a wide range of function types to see which one is best: `mReg`  
```{r regression, fig.show='hold', echo=-(1:2)}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
options(digits=5)
a <- 1:30   ; b <- a/2.345+rnorm(30,0,3)
linReg(a,b, main="linReg, circle, addAlpha")

circle(12,3, r=5, col=addAlpha("darkgreen"), border="blue", lwd=3)

x <- c(1.3, 1.6, 2.1, 2.9, 4.4, 5.7, 6.6, 8.3, 8.6, 9.5)
y <- c(8.6, 7.9, 6.6, 5.6, 4.3, 3.7, 3.2, 2.5, 2.5, 2.2)
mReg(x,y, main="mReg")[,c(2,3,5:6)]
```

Table with numbers and corresponding color: `tableColVal`
```{r tableColVal, echo=-1, fig.height=3.5, fig.width=5.5}
par(mar=c(0,0,1,0))
tableColVal(as.matrix(eurodist)[1:15,1:5], nameswidth=0.25)
```

Climate diagram: `climateGraph`
```{r climgraph, echo=-1, fig.height=3.5, fig.width=5.5}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
climateGraph(temp=c(-9.3,-8.2,-2.8,6.3,13.4,16.8,18.4,17,11.7,5.6,-1,-5.9),
             rain=c(46,46,36,30,31,21,26,57,76,85,59,46))
```

[TOC](#header)

## Dataframe Operations
```{r df}
# Convert list with vectors of unequal length to one single data.frame: l2df
eglist <- list(BB=c(6,9,2,6), KA=1:8, JE=c(-3,2) )
eglist
l2df(eglist)  # names are even kept

# add rows to a data.frame: addRows, insertRows
MYDF <- data.frame(A=5:3, B=2:4)
addRows(MYDF, 3)
insertRows(MYDF, 2, 10:11)

# Order rows in a dataframe: sortDF
sortDF(USArrests[USArrests$Murder>14,], "Assault", decreasing=TRUE)

# truth table to test logical expressions: TFtest
TFtest(!a & !b, a&b, !(a&b))

# Head and tail at the same time: headtail (exception from lowerCamelCasing)  
headtail(iris, n=3, na=FALSE)
```

[TOC](#header)

## Graphics

Color palettes: `seqPal`, `divPal`, `showPal`  
Plot simulation results as result ranges: `quantileBands`, `ciBand`  
```{r showPal, fig.show='hold'}
showPal(cex=0.55)

neff <- t(replicate(n=300, sapply(1:200, function(nn) max(rnorm(nn)))   ))
qB <- quantileBands(neff, x=1:200, smooth=7)
```

Quickly plot distributions by just specifying parameters: `normPlot`, `betaPlot`
```{r distrplots, fig.show='hold', echo=-1}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
normPlot(mean=81.7, sd=11.45)
betaPlot(shape1=1.5, shape2=6) 
```

Compare Beta distribution parameter effects: `betaPlotComp`  
```{r betacomp, echo=-1, fig.height=4.5, fig.width=5.5}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
betaPlotComp()
```

Set ylim so that it does not extend below zero: `lim0`
```{r lim0, fig.show='hold', echo=-1}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
val <- c(3.2, 1.8, 4.5, 8.2, 0.1, 2.9) # just some numbers
plot(val) # axes are extended by 4\% automatically, if xaxs="r"
plot(val, ylim=lim0(val), las=1) # you don't even have to set yaxs="i" ;-)
```

Histogram with bars drawn horizontally: `horizHist`  
Histograms for dataset split into categories: `groupHist`  
(Uses `panelDim` to compute layout of panels passed to `par(mfrow)`)
```{r horizHist, fig.show='hold', echo=-1}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
ExampleData <- rnorm(200,13,5)
hpos <- horizHist(ExampleData, col=4)
abline(h=hpos(11), col=2, lwd=2)

groupHist(chickwts, "weight", "feed", col=2, unit="gr_6")
# drop the horsebean, feed those chicks with sunflower seeds (unless you like small chicken)
```

A few interactive things (not shown als Graphs)  
Zoom into graphics: `pointZooom`  
Horizontal and Vertial line at point clicked on: `locLine`  
Transformation from linear to logarithmic axis: `linLogHist`, `linLogTrans`  
```{r pointZoom, eval=FALSE}
a <- rnorm(90); b <- rexp(90)
dev.new(record=TRUE) # turn recording on
plot(a,b, las=1)
pointZoom(a,b) # scroll through the plots (Pg Up and Pg Dn) to unzoom again.

locLine()

x <- rlnorm(700, m=3)
dev.new(record=TRUE) # scroll through the plots (Pg Up and Pg Dn)...
linLogHist(x, xlab="ddd", breaks=30, yaxt="n", freq=FALSE)
```
 

Moving average with overlapping windows: `movAv`, `movAvLines`  
Funnel plot for proportional Data: `funnelPlot`  
```{r movAvLines, fig.show='hold', echo=-(1:2), warning=FALSE}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
set.seed(42); a <- cumsum(rnorm(100))
plot(a, type="l", pch=16, las=1)
lines(movAv(a), col=2, lwd=3)
movAvLines(a, lwd=3, plot=FALSE)

X <- c(2, 224,  54,  505, 1,  5, 236,  92,  3, 0) # successful events
N <- c(2, 400, 100, 1000, 2, 10, 500, 200, 10, 2) # possible succeses
funnelPlot(X,N, letters[1:10])
```
  
Get nice values and labels to write at logarithmic axes: `logVals`, `logAxis`  
Label time axis in date-regular intervals: `monthLabs`, `monthAxis`
```{r axes, fig.show='hold', echo=-(1:2)}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
set.seed(42)
exdat <- 10^runif(50, -1, 2)
plot(exdat, log="y", yaxt="n")
logAxis(side=2) # invisibly returns values and labels
points(exdat, pch=16)

plot(as.Date("2013-10-25")+sort(sample(0:100, 30)), 
     cumsum(rnorm(30)), type="o", xaxt="n", ann=F)
dummy <- monthAxis(side=1, n=4) # tries to find suitable labeling intervall
str(dummy)
```


[TOC](#header)

## Hydrology

Extreme value Statistics (e.g. for flood risk estimation):
moved to https://github.com/brry/extremeStat

```{r hydro, echo=-1}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))
# superposition of precipitation to simulate Q from P: superPos
N <- c(9,5,2,14,1,3) # [mm/hour]
UH <- c(0.1, 0.4, 0.3, 0.1, 0.1) # [1/h]
superPos(N, UH)

# calculate continuous UH with given n and k: unitHydrograph
plot(0:40, unitHydrograph(n=2,  k=3, t=0:40), type="l")

# Nash-Sutcliffe efficiency: nse
QSIM <- lsc(PREC, QOBS, area=10, returnsim=TRUE, plot=FALSE)
nse(QOBS, QSIM)

# Root Mean Squared Error, e.g. to be minimized: rmse
rmse(QOBS, QSIM)

# R squared (coefficient of determination): rsquare
rsquare(QOBS, QSIM)
```


[TOC](#header)

## Spatial
```{r spatial, echo=-1, fig.height=3.5, fig.width=5.5}
par(mar=c(3.2,3.2,1.5,0.7), mgp=c(2.1,0.7,0))

# arrange points randomly with minimal distance to each other: randomPoints
rp <- randomPoints(xmin=5,xmax=15, ymin=20,ymax=30, number=25, mindist=1)

# distance between two points on a plane: distance
A <- c(3,  9,-1)  ;  B <- c(7, -2, 4)
plot(A,B); points(3,5, col=2, pch=16); segments(3,5, A,B)
distance(A,B, 3,5)

# calculate Area of a planar triangle: triangleArea
a <- c(1,5.387965,9); b <- c(1,1,5)
plot(a[c(1:3,1)], b[c(1:3,1)], type="l", asp=1)
abline(v=1:9, h=1:5, col=8,lty=2)
triangleArea(a,b)
```

Zooming in Shapefiles in my package https://github.com/brry/shapeInteractive


[TOC](#header)

## Miscellaneous
```{r misc}
# paste character strings with collapse: pastec
listoferrors <- c("filetype", "header", "nonemptyline")
message("The following entities were corrupted:\n", pastec(listoferrors))

# remove leading and trailing white space: removeSpace
s <- c("space at end     ", "  white at begin", "  both ", " special ^  ")
removeSpace(s)

# calculate confidence interval (and other EDA values): ci, cie
ci(  c(5:8,3,14) )

# sequence given by range or vector of values: seqR
seqR(range=c(12,6), by=-2)
seqR(rnorm(20), len=7)

# Rescale values to another range: rescale
rescale(10:15, from=200, to=135)

# Show memory size of the biggest objects in MB: lsMem
lsMem(n=5)

# extract pdf link from google search result url: googleLink2pdf
Link <- paste0("http://www.google.de/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1",
        "&cad=rja&sqi=2&ved=0CDIQFjAA&url=http%3A%2F%2Fcran.r-project.org",
        "%2Fdoc%2Fmanuals%2FR-intro.pdf&ei=Nyl4UfHeOIXCswa6pIC4CA",
        "&usg=AFQjCNGejDwPlor4togQZmQEQv72cK9z8A&bvm=bv.45580626,d.Yms")
googleLink2pdf(Link)

# Create a number of 999 strings with spaces for reading files: na9
na9()[c(1:4,13,30)]

```

A few things not executed for this document:
```{r misc_non, eval=FALSE}
# Separate lists with arguments for functions: owa
?owa # the example section has a good - wait for it - example!

# install.package and require in one single function: require2
require2(ada)

# Write a file with a Roxygen-compatible function structure, 
# making it easy to add new functions to the package: createFun
createFun(myNewFunction, package="extremeStat", path="S:/Dropbox")

# Open the source code of a function on github: funSource
funSource("smoothLines")

# Install a package from github without dependencies: instGit
instGit("brry/shapeInteractive")


# concatenate textfiles contents unchanged into one file: combineFiles
# see also: compareFiles, dupes
cat("This is Sparta.\nKicking your face.", file="BujakashaBerry1.txt")
cat("Chuck Norris will roundhousekick you.", file="BujakashaBerry2.txt")
combineFiles(inFiles=paste0("BujakashaBerry", 1:2, ".txt"),
                 outFile="BujakashaBerry3.txt")
readLines("BujakashaBerry3.txt")
unlink(paste0("BujakashaBerry", 1:3, ".txt"))

# wish neRds a happy new year: yearSample
yearSample(2016)
# Have a nerdy
set.seed(12353); sample(0:9,4,T)

# generate name from "random" sample: nameSample
nameSample("berry")
```
```{r misc2}
# Kind regards from
set.seed(1248272); paste(sample(letters,5,rep=T),collapse='')
```


[TOC](#header)

Explore the other possibilities of the package by reading the function help files.  
Any Feedback on this package (or this vignette) is very welcome via github or <berry-b@gmx.de>!
